Phoneme surprisal

### 2019-06-26 surprisal/letter_surprisal.Rmd
providence unigrams, palasis and wagner letter surprisal

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(childesr)

@@

```{r set up corpus}
prov_utterances <- get_utterances(corpus = "Providence")

@@

```{r plot results}
letter_unigrams %>%
  filter(word_length %in% 3:11) %>%
  ggplot(aes(x = letter_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(childesr)
library(here)
library(feather)

@@

```{r set up corpus}
prov_utterances <- get_utterances(corpus = "Palasis")

@@

```{r plot results}
letter_unigrams <- letter_unigrams %>% mutate(gram = "Unigram") 
letter_bigrams <- letter_bigrams %>% mutate(gram = "Bigram") 
letter_trigrams <- letter_trigrams %>% mutate(gram = "Trigram") 
letter_unigrams %>% bind_rows(letter_bigrams, letter_trigrams) %>% write_feather(here("Data/palasis_letters.feather"))

letter_unigrams %>%
  filter(word_length %in% 3:11) %>%
  ggplot(aes(x = letter_order, y = mean_s)) + 
    facet_wrap(~word_length) + 
    geom_point() +
    geom_line() + 
    xlab("Letter Position") + 
    ylab("Mean surprisal")
```

### 2019-06-26 surprisal/phoneme_surprisal.Rmd
providence unigrams, palasis and wagner letter surprisal

@@

```{r show leave-out-surprisal}
split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))

>>>>

```{r childes unigrams}
phone_probs <- adult_phonemes %>% 
  count(phone) %>%
  ungroup() %>%
  mutate(p = n/sum(n))

phone_unigrams <- adult_phonemes %>% 
  left_join(phone_probs) %>%
  mutate(s = -log(p)) %>%
  group_by(word_length, phone_order) %>%
  summarise(mean_s = mean(s))
```

```{r childes bigrams}
bigram_probs <- adult_phonemes %>% 
  group_by(utterance_id) %>%
  mutate(lag_phone = lag(phone)) %>%
  group_by(lag_phone, phone) %>%
  count() %>%
  filter(!is.na(lag_phone)) %>%
  ungroup() %>%
  left_join(phone_probs, by = c("lag_phone" = "phone")) %>%
  mutate(cond_p = n.x / n.y) %>%
  select(lag_phone, phone, cond_p, n.x) %>%
  rename(n = n.x)

phone_bigrams <- adult_phonemes %>%
  group_by(utterance_id, word_id) %>%
  mutate(lag_phone = lag(phone)) %>%
  ungroup() %>%
  left_join(bigram_probs, by = c("lag_phone" = "lag_phone", "phone" = "phone")) %>%
  left_join(phone_probs, by = c("phone" = "phone")) %>%
  mutate(s = ifelse(is.na(lag_phone), -log(p), -log(cond_p))) %>%
  group_by(word_length, phone_order) %>%
  summarise(mean_s = mean(s)) 
```

```{r childes trigrams}
trigram_probs <- adult_phonemes %>%
  group_by(utterance_id, word_id) %>%
  mutate(lag_phone1 = lag(phone)) %>%
  mutate(lag_phone2 = lag(lag_phone1)) %>%
  group_by(lag_phone2, lag_phone1, phone) %>%
  count() %>%
  filter(!is.na(lag_phone1), !is.na(lag_phone2)) %>%
  ungroup() %>%
  left_join(bigram_probs, by = c("lag_phone2" = "lag_phone", "lag_phone1" = "phone")) %>%
  mutate(tri_cond_p = n.x / n.y) %>%
  select(-cond_p, -n.y) %>%
  rename(n = n.x)

phone_trigrams <- adult_phonemes %>%
  group_by(utterance_id, word_id) %>%
  mutate(lag_phone1 = lag(phone)) %>%
  mutate(lag_phone2 = lag(lag_phone1)) %>%
  left_join(trigram_probs, by = c("lag_phone2" = "lag_phone2", "lag_phone1" = "lag_phone1", "phone" = "phone")) %>% 
  left_join(bigram_probs, by = c("lag_phone1" = "lag_phone", "phone" = "phone")) %>% 
  left_join(phone_probs, by = c("phone" = "phone")) %>%
  mutate(s = ifelse(is.na(lag_phone2), # check if it's not the first word
                    ifelse(is.na(lag_phone1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(word_length, phone_order) %>%
  summarise(mean_s = mean(s)) %>% 
  filter(phone_order <= word_length)
```

@@

```{r show leave-out-entropy}
split_lengths <- adult_phonemes %>%
  group_by(word_length) %>%
  nest() %>%
  filter(word_length > 1) %>%
  filter(word_length %in% 1:6 ) %>%
  mutate(entropy = map(data, hold_out_entropy))
